---
title: "Bollywood Movies"
output: html_document
date: "2025-02-20"
names: HS, AK, IS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(ggplot2)
library(RColorBrewer)
```


```{r}
options(scipen=999) #turning off scientific notation
```


Load Data:
```{r}
movies1 <- read_csv("bollywood.csv") %>% 
  clean_names() %>% 
  rename(title=movie_name) %>% 
  mutate(across(where(is.character), tolower)) #changing all characters to lowercase
```

```{r}
movies2 <- read_csv("bollywood_hits.csv") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower))
```

Join Data:
```{r}
movie_join <- inner_join(movies1, movies2, by="title")
```



Cleaning the Data
```{r}
bollywood <- movie_join %>% 
  select(-genre.y, -actors, -directors, -imdb_id) %>% #removing duplicate columns
  rename(genre=genre.x) %>% 
  separate(release_date, into=c("date", "month", "year"), sep=" ") %>% #separate the date column
  replace_with_na_all(condition = ~.x == "n/a") %>% 
  select(-release_year) #removing duplicate year column
  
```

Assessing data:

```{r}
summary(bollywood)
```

```{r}
miss_var_summary(bollywood)
```





##Change Over Time

1. How does the number of movies released change over the years?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  ggplot(aes(x=as.factor(year)))+
  geom_bar(fill="skyblue", color="black")+
  labs(title="Movies per Year", x="Year", y="Number Movies")+
  theme(plot.title = element_text(hjust=0.5, size=15))
```
Overall general increase in number of movies
2011 has highest number of movies and 2006 has the lowest


2. Could the drop in movies in 2006 be related to the number of new directors?
```{r}
bollywood %>% 
  filter(new_director=="yes", year!="NA") %>% 
  group_by(year) %>% 
  summarize(n_uniq=n_distinct(director))%>% 
  ggplot(aes(x=as.factor(year), y=n_uniq))+
  geom_col(fill="skyblue", color="black")+
  labs(title="New Directors per Year", x="Year", y="Number of New Directors")+
  theme(plot.title = element_text(hjust=0.5, size=15))
```
2006 is the year with the fewest directors entering the field; corresponds to fewest movies released
2011 is the year with the highest new directors entering the field; corresponds to most movies released


3. How does revenue (movie success) change over time?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  group_by(year) %>%
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev))+
  geom_col(fill="skyblue", color="black")+
  labs(title="Average Revenue over Time", x="Year", y="Average Revenue")+
  theme(plot.title=element_text(hjust=0.5, size=15))
```
Average revenue is generally increasing over time but there is a drop from 2009-2011


4. How does the distribution of revenue change over time?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  ggplot(aes(x=year, y=revenue_inr))+
  geom_boxplot(fill="skyblue")+
  labs(title="Revenue per Year", x="Year", y="Revenue")+
  theme(plot.title=element_text(hjust=0.5, size=15))
  
```
Number of high outliers increase over time
Outliers are extremely high from 2012-2014


5. From 2012-2014, how does the revenue range vary by genre?
```{r}
bollywood %>% 
  filter(year %in% c(2012, 2013, 2014)) %>% 
  filter(!genre %in% c("adult", "suspense","fantasy")) %>% 
  ggplot(aes(x=genre, y=revenue_inr/10000, fill=genre))+
  geom_boxplot()+
  facet_wrap(~year)+
  scale_fill_brewer(palette="RdYlGn")+
  coord_flip()+
  theme(axis.text.x=element_text(angle=20))+
  labs(title="Revenue Range over Time", y="Revenue x 10^-3 in Rupees", x="Genre")
  
```
Masala movies revenue increased significantly across the 3 years
Range in revenue of action movies increased a lot in 2014

6. How did the number of masala movies change from 2012-2014
```{r}
bollywood %>% 
  filter(year %in% c(2012, 2013, 2014)) %>% 
  filter(genre=="masala") %>% 
  group_by(year) %>% 
  summarize(n_masala=n())
```
Only 2 masala movies in 2013 and 2014 but they did really well

7. What were the names of these movies?
```{r}
bollywood %>% 
  filter(year %in% c(2013, 2014)) %>% 
  filter(genre=="masala") %>% 
  select(title, lead_star, director, year, revenue_inr)
```


8.  How did the success of the lead actors involved in the masala movies from Q7 change over time?
```{r}
bollywood %>% 
  filter(lead_star=="ajay devgn" | lead_star=="shahrukh khan"| lead_star=="ranveer singh") %>% 
  group_by(year, lead_star) %>% 
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev, group=lead_star, color=lead_star))+
  geom_point()+
  geom_line()+
  labs(title="Avg Movie Revenue Over Time", x="Year", y="Average Revenue", color="Actor")+
  theme(plot.title=element_text(hjust=0.5, size=15))
  
```
Within 2005-2014:
ajay devgun and shah rukh khan: their success peaked in 2014
ranveer singh had just entered the industry in 2013

9. How did the success of the directors involved in the masala movies from Q7 change over time?
```{r}
bollywood %>% 
  filter(director=="sajid khan" | director=="ali abbas zafar"| director=="farah khan" | director=="rohit shetty") %>% 
  group_by(year, director) %>% 
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev, group=director, color=director))+
  geom_point()+
  geom_line()+
  labs(title="Avg Movie Revenue Over Time", x="Year", y="Average Revenue", color="director")+
  theme(plot.title=element_text(hjust=0.5, size=15))
  
```



#Hits vs. Flops





#Old vs. New Bollywood




