---
title: "Bollywood Movies"
output: html_document
date: "2025-02-20"
names: HS, AK, IS
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(ggplot2)
library(RColorBrewer)
```


```{r}
options(scipen=999) #turning off scientific notation
```


Load Data:
```{r}
movies1 <- read_csv("movies1.csv") %>% 
  clean_names() %>% 
  rename(title=movie_name) %>% 
  mutate(across(where(is.character), tolower)) #changing all characters to lowercase
```

```{r}
movies2 <- read_csv("bollywood_hits.csv") %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower))
```

Join Data:
```{r}
movie_join <- inner_join(movies1, movies2, by="title")
```



Cleaning the Data
```{r}
bollywood <- movie_join %>% 
  select(-genre.y, -actors, -directors, -imdb_id) %>% #removing duplicate columns
  rename(genre=genre.x) %>% 
  separate(release_date, into=c("date", "month", "year"), sep=" ") %>% #separate the date column
  replace_with_na_all(condition = ~.x == "n/a") %>% 
  select(-release_year) #removing duplicate year column
  
```

Assessing data:

```{r}
summary(bollywood)
```

```{r}
miss_var_summary(bollywood)
```





##Change Over Time

1. How does the number of movies released change over the years?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  ggplot(aes(x=as.factor(year)))+
  geom_bar(fill="skyblue", color="black")+
  labs(title="Movies per Year", x="Year", y="Number Movies")+
  theme(plot.title = element_text(hjust=0.5, size=15))
```
2011 has highest number of movies and 2006 has the lowest


2. Could the drop in movies in 2006 be related to the number of new directors?
```{r}
bollywood %>% 
  filter(new_director=="yes", year!="NA") %>% 
  group_by(year) %>% 
  summarize(n_uniq=n_distinct(director))%>% 
  ggplot(aes(x=as.factor(year), y=n_uniq))+
  geom_col(fill="skyblue", color="black")+
  labs(title="New Directors per Year", x="Year", y="Number of New Directors")+
  theme(plot.title = element_text(hjust=0.5, size=15))
```
2006 is the year with the fewest directors entering the field; corresponds to fewest movies released
2011 is the year with the highest new directors entering the field; corresponds to most movies released
Number of new directors could factor into the number of movies released per year


3. How does revenue change over time?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  group_by(year) %>%
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev))+
  geom_col(fill="skyblue", color="black")+
  labs(title="Average Revenue over Time", x="Year", y="Average Revenue in Rupees")+
  theme(plot.title=element_text(hjust=0.5, size=15))
```
Average revenue is generally increasing over time (growth of bollywood industry)


4. How does the distribution of revenue change over time?
```{r}
bollywood %>% 
  filter(year != "NA") %>% 
  ggplot(aes(x=year, y=revenue_inr, fill=year))+
  geom_boxplot()+
  scale_fill_brewer(palette="RdYlGn")+
  labs(title="Revenue per Year", x="Year", y="Revenue (INR)")+
  theme(plot.title=element_text(hjust=0.5, size=15))
```
Value of high outliers increasing over time
Outliers are extremely high from 2012-2014





5. From 2012-2014, how does the revenue range vary by genre?
```{r}
bollywood %>% 
  filter(year %in% c(2012, 2013, 2014)) %>% 
  filter(!genre %in% c("adult", "suspense","fantasy")) %>% #removing some genres that only apply to very few movies with low medians to make graph cleaner
  ggplot(aes(x=genre, y=revenue_inr/10000, fill=genre))+
  geom_boxplot()+
  facet_wrap(~year)+
  scale_fill_brewer(palette="RdYlGn")+
  coord_flip()+
  theme(axis.text.x=element_text(angle=20))+
  labs(title="Revenue Range over Time", y="Revenue x 10^-3 (Rupees)", x="Genre")
  
```
Masala and action movies have higher medians than other genres (these are probably contributing to high outliers from Q4)
Masala movie boxplot shifted to the right significantly from 2012 to 2013 (increasing success of the movies)


6. Did the number of masala movies change from 2012-2014?
```{r}
bollywood %>% 
  filter(year %in% c(2012, 2013, 2014)) %>% 
  filter(genre=="masala") %>% 
  group_by(year) %>% 
  summarize(n_masala=n())
```
Number of masala movies remained low
Only 2 masala movies in 2013 and 2014 but they did really well according to boxplot from Q5


7. What were the names of these movies? Who were the actors and directors involved?
```{r}
bollywood %>% 
  filter(year %in% c(2013, 2014)) %>% 
  filter(genre=="masala") %>% 
  select(title, lead_star, director, year, revenue_inr) %>% 
  arrange(year)
```


8.  How did the success of the lead actors involved in the masala movies from Q7 change over time?
```{r}
bollywood %>% 
  filter(lead_star=="ajay devgn" | lead_star=="shahrukh khan"| lead_star=="ranveer singh") %>% 
  group_by(year, lead_star) %>% 
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev, group=lead_star, color=lead_star))+
  geom_point()+
  geom_line()+
  labs(title="Movie Revenue Over Time", x="Year", y="Average Revenue (INR)", color="Actor")+
  theme(plot.title=element_text(hjust=0.5, size=15))
  
```
Shahrukh khan and ajay devgn show impressive growth over time, reach peak in 2014
Ranveer singh is newer to the industry

9. How did the success of the directors involved in the masala movies from Q7 change over time?
```{r}
bollywood %>% 
  filter(director=="sajid khan" | director=="ali abbas zafar"| director=="farah khan" | director=="rohit shetty") %>% 
  group_by(year, director) %>% 
  summarize(avg_rev=mean(revenue_inr)) %>% 
  ggplot(aes(x=year, y=avg_rev, group=director, color=director))+
  geom_point(size=2)+
  geom_line()+
  labs(title="Movie Revenue Over Time", x="Year", y="Average Revenue (INR)", color="director")+
  theme(plot.title=element_text(hjust=0.5, size=15))
  
```
2013-2014 were the most successful year for all the directors except Sajid Khan
Rohit shetty's masala movie was important step in his career
Farah khan had huge jump from 2010-2014 so the masala movie she produced was also huge step in her career




#Old vs. New Bollywood

```{r}
names(bollywood)
```

```{r}
write.table(bollywood, "bollywood.txt", sep = "\t", row.names = FALSE)
```


```{r}
write.csv(bollywood, "bollywood.csv", row.names = FALSE)
```

```{r}
write.csv(bollywood_eras, "bollywood_eras.csv", row.names = FALSE)
```

1. How many movies belong to "Old Bollywood" vs "New Bollywood"
```{r}
bollywood_era <- bollywood %>% 
  mutate(bollywood_era = case_when(
    year >= 2005 & year <= 2009 ~ "Old Bollywood",
    year >= 2010 & year <= 2014 ~ "New Bollywood",
    TRUE ~ "Other"
  )) %>% 
  group_by(bollywood_era) %>% 
  summarise(count=n())
```


```{r}
bollywood_era %>%
  count(bollywood_era) %>%
  ggplot(aes(x = bollywood_era, y = n, fill = bollywood_era)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(title = "Number of Movies in Old vs. New Bollywood", x = "Bollywood Era", y = "Number of Movies") +
  theme_minimal() +
  scale_fill_manual(values = c("Old Bollywood" = "blue", "New Bollywood" = "red")) +
  theme(legend.position = "none", text = element_text(size = 12))
```
```{r}
bollywood_eras <- bollywood %>% 
  mutate(era=case_when(year<=2009 ~"old",
                       year>2009 ~"new"))
```

2. How does the average revenue of movies compare between the old and new Bollywood eras?

```{r}
bollywood_erasclean <- bollywood_eras %>% 
  filter(!is.na(revenue_inr), !is.na(era))
```

```{r}
revenue_comparison <- bollywood_erasclean %>%
    group_by(era) %>%
    summarise(avg_revenue = mean(revenue_inr, na.rm = TRUE))
```

```{r}
ggplot(revenue_comparison, aes(x = era, y = avg_revenue, fill = era)) +
    geom_bar(stat = "identity", color = "black", width = 0.5) +
    scale_fill_manual(values = c("old" = "steelblue", "new" = "red")) +
    labs(title = "Average Revenue Comparison: Old vs. New Bollywood",
         x = "Era",
         y = "Average Revenue (INR)") +
    theme_minimal() +
    theme(text = element_text(size = 14), legend.position = "none")
```




3. What are the most common genres in old and new Bollywood Eras?
```{r}
bollywood_erasclean <- bollywood_eras %>% 
  filter(!is.na(genre), !is.na(era))
```

```{r}
genre_count <- bollywood_erasclean %>%
    group_by(era, genre, .groups='keep') %>%
    summarise(count = n()) %>%
    arrange(desc(count))
```

4. How do hit and flop movies differ in the old and new Bollywood eras?
```{r}
bollywood_erasclean <- bollywood_eras %>% 
  filter(!is.na(hit_flop.x), !is.na(era))
```





1. How does the revenue distribution compare between Old and New Bollywood eras?
```{r}
bollywood_eras %>%
  filter(!is.na(revenue_inr), !is.na(era)) %>%
  ggplot(aes(x=era, y=revenue_inr, fill=era)) +
  geom_boxplot(outlier.color="red", outlier.shape=16, alpha=0.7) +
  theme_minimal() +
  scale_fill_manual(values=c("#FF5733", "#33FF57")) +
  labs(title="Revenue Distribution by Era", subtitle="Comparing old and new Bollywood movies",
       x="Era", y="Revenue (INR)", fill="Era")
```

This plot compares the revenue distribution between old and new Bollywood movies. The median revenue, spread, and presence of outliers help us understand whether newer films tend to have higher earnings or if both eras have similar revenue patterns. Since the newer era shows a higher median, it suggests Bollywood films have become more commercially successful according to box office.


2. Which genres are more dominant in Old vs New Bollywood?
```{r}
bollywood_eras %>% 
  filter(!is.na(genre), !is.na(era)) %>%
  group_by(genre, era) %>%
  summarise(count=n(), .groups='keep') %>%
  ggplot(aes(x=reorder(genre, count), y=count, fill=era)) +
  geom_col(position="dodge") +
  theme_minimal() +
  coord_flip() +
  scale_fill_manual(values=c("#FF5733", "#33FF57")) +
  labs(title="Number of Movies per Genre", subtitle="Comparing old and new Bollywood genres",
       x="Genre", y="Count", fill="Era")
```
This plot illustrates the number of movies per genre for both old and new Bollywood. If certain genres are significantly more popular in one era compared to the other, it could suggest a shift in audience preferences over time. For instance, action films might dominate the newer era, while romance or drama could be more prevalent in older Bollywood.


4. How does the number of screen relate to revenue for ecah era?
```{r}
bollywood_eras %>% 
  filter(!is.na(number_of_screens), !is.na(revenue_inr), !is.na(era)) %>%
  ggplot(aes(x=number_of_screens, y=revenue_inr, color=era)) +
  geom_point(size=2, alpha=0.6) +
  theme_minimal() +
  scale_color_manual(values=c("#FF5733", "#33FF57")) +
  labs(title="Number of Screens vs Revenue", subtitle="Does wider release mean higher revenue?",
       x="Number of Screens", y="Revenue (INR)", color="Era")
```
This graph helps us understand if releasing a movie on more screens leads to higher revenue. If we see an upward trend, it means movies released on more screens generally make more money and vice versa for the downward trend. So the plot basically shows the more places a movie is show, the more money it makes.

5. Who are the most popular directors in Old and New Bollywood?
```{r}
popular_directors <- bollywood_eras %>%
  filter(!is.na(director), !is.na(era)) %>%
  group_by(director, era) %>%
  summarise(movie_count = n(), .groups = 'keep') %>%
  arrange(desc(movie_count))
```

```{r}
top_directors <- popular_directors %>%
  group_by(era) %>%
  top_n(10, movie_count)
```

```{r}
ggplot(top_directors, aes(x = reorder(director, movie_count), y = movie_count, fill = era)) +
  geom_col(alpha = 0.8, show.legend = TRUE) +
  coord_flip() +  
  theme_minimal() +
  scale_fill_manual(values = c("old" = "#FF5733", "new" = "#33FF57")) +
  labs(
    title = "Top 10 Most Popular Directors in Old and New Bollywood",
    subtitle = "Based on the number of movies directed",
    x = "Director",
    y = "Number of Movies",
    fill = "Era"
  )
```
There are some actors that didn't direct any of the older movies, some that directed only older movies, and some that directed movies in both eras.

5. How do the eras compare when a new_actor is casted in a movie?

```{r}
new_actor_data <- bollywood_eras %>%
  drop_na(new_actor, era) %>%
  group_by(era, new_actor, .groups = 'keep') %>%
  summarise(movie_count = n())

ggplot(new_actor_data, aes(x = era, y = movie_count, fill = new_actor)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Comparison of New Actors Cast in Movies Across Eras",
    x = "Era",
    y = "Number of Movies",
    fill = "New Actor Cast"
  )
```
There are less new actors in the old era.

6. How did the revenue compare in both eras when Shah Rukh Khan was cast?
```{r}
srk_revenue_data <- bollywood_eras %>%
  filter(lead_star == "shahrukh khan") %>%
  drop_na(revenue_inr, era) %>%
  group_by(era, .groups = 'keep') %>%
  summarise(avg_revenue = mean(revenue_inr))

ggplot(srk_revenue_data, aes(x = era, y = avg_revenue, fill = era)) +
  geom_col() +
  theme_minimal() +
  labs(
    title = "Average Revenue of Movies Starring Shah Rukh Khan Across Eras",
    x = "Era",
    y = "Average Revenue (INR)"
  )
```

7. Most popular genre during the holiday release period in both eras
```{r}
genre_popularity_data <- bollywood_eras %>%
  filter(release_period == "holiday") %>%
  drop_na(genre, era) %>%
  group_by(era, genre, .groups = 'keep') %>%
  summarise(movie_count = n())

ggplot(genre_popularity_data, aes(x = genre, y = movie_count, fill = era)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 60))
  theme_minimal() +
  labs(
    title = "Most Popular Genre During Holiday Releases Across Eras",
    x = "Genre",
    y = "Number of Movies",
    fill = "Era"
  )
```
everybody likes drama

8. Out of these music directors, which one had the highest hit_flop average: a.r. rehman, pritam, vishal - shekhar, anu malik.
```{r}
music_director_hit_flop <- bollywood_eras %>%
  filter(music_director %in% c("a.r. rehman", "pritam", "vishal - shekhar", "anu malik")) %>%
  group_by(era, music_director, .groups = 'keep') %>%
  summarise(hit_flop_avg = mean(hit_flop, na.rm = TRUE))

ggplot(music_director_hit_flop, aes(x = reorder(music_director, hit_flop_avg), y = hit_flop_avg, fill = era)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Highest Hit-Flop Average by Music Directors Across Eras",
    x = "Music Director",
    y = "Hit-Flop Average",
    fill = "Era"
  )
```

8. How does the revenue distribution by genre differ between the two eras? (Faceted Plot)
```{r}
revenue_genre_era <- bollywood_eras %>%
  drop_na(genre, revenue_inr, era) %>%
  group_by(era, genre, .groups = 'keep') %>%
  summarise(avg_revenue = mean(revenue_inr))

ggplot(revenue_genre_era, aes(x = genre, y = avg_revenue, fill = genre)) +
  geom_col() +
  facet_wrap(~era) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Revenue Distribution by Genre Across Eras",
    x = "Genre",
    y = "Average Revenue (INR)",
    fill = "Genre"
  )
```



#Hits vs. Flops

1. How many hits and flops?
```{r}
bollywood <- bollywood %>%
  mutate(success = case_when(
    hit_flop < 7 ~ "flop", 
    hit_flop >= 7 ~ "hit")) 
bollywood %>% 
  group_by(success) %>%
  summarize(hit_count = n()) %>%
    ungroup()
```
699 hits and 42 flops

2. Compare hits and flops by whether they are remakes or not
```{r}
bollywood %>%
  filter(!is.na(whether_remake)) %>%
  group_by(whether_remake, success) %>%
  summarize(count = n(), .groups = "drop")

```

3. How do the number of hit and flop movies compare for remakes versus non-remakes in the Bollywood dataset?
```{r}
bollywood %>%
  filter(!is.na(whether_remake)) %>%
  group_by(whether_remake, success) %>%
  summarize(count = n(), .groups = "drop") %>% 
  ggplot(aes(x=success, y=count, fill=whether_remake))+
  geom_col()+
  labs(title="Comparison of Hits and Flops by Whether Remake")

```

4. Compare hits and flops by the type genre. Create the appropriate plot
```{r}
genre_summary <- bollywood %>%
  group_by(genre, hit_flop) %>%
  summarize(count = n(), .groups = "drop")

```

```{r}
ggplot(genre_summary, aes(x = genre, y = count, fill = hit_flop)) +
  geom_bar(stat = "identity", position = "dodge") +  # Dodge position to show bars side by side
  labs(
    title = "Comparison of Hits and Flops by Genre",
    x = "Genre",
    y = "Number of Movies",
    fill = "Hit/Flop Status"
  ) +
  theme_minimal() +
  coord_flip()

```

5.  Compare hits and flops by the following actors: aamir bashir, aasad mirza,abhay deol and abhishek bachchan. Create a table.
```{r}
actor_summary <- bollywood %>%
  filter(lead_star %in% c("aamir bashir", "aasad mirza", "abhay deol", "abhishek bachchan"))%>%
  group_by(hit_flop, lead_star) %>%
  summarize(count = n(), .groups = "keep") %>% 
  arrange(desc(count))
actor_summary

```


6.  Does release period relate to the success of hits and failure of flops? Create a table.

```{r}
bollywood <- bollywood %>%
  mutate(release_period = as.factor(release_period))
success_by_period <- bollywood %>%
  group_by(release_period, hit_flop) %>%
  summarize(count = n(), .groups = "drop")
success_by_period

```

7. How do the number of hit and flop movies vary for Abhay Deol and Aishwarya Rai in the Bollywood dataset? 
```{r}
actor_summary <- bollywood %>%
  filter(lead_star %in% c("abhay deol", "aishwarya rai")) %>%
  group_by(hit_flop, lead_star) %>%
  summarize(count = n(), .groups = "keep")

```

```{r}
ggplot(actor_summary, aes(x = hit_flop, y = count, fill = lead_star)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Lead Star", y = "Movie Count", fill = "Lead Star") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


8. 

9. What is the average number of hit and flop movies for each release period among the selected Bollywood lead stars?
```{r}
success_by_period <- bollywood %>%
  filter(lead_star %in% c("aamir bashir", "aasad mirza", "abhay deol", "abhishek bachchan")) %>%
  group_by(release_period, hit_flop) %>%
  summarize(count = n(), .groups = "drop")

```

```{r}
average_success_by_period <- success_by_period %>%
  group_by(release_period) %>%
  summarize(average_count = mean(count), .groups = "drop")
average_success_by_period
```




## Shiny App

```{r}
by_genre <- bollywood %>% 
  group_by(genre) %>% 
  summarize(avg_hit_rating=mean(hit_flop), avg_num_screens=mean(number_of_screens), avg_revenue_inr=mean(revenue_inr), avg_budget_inr=mean(budget_inr))
```


```{r}
library(shiny)

ui <- fluidPage(titlePanel("Bollywood Genres"),
  selectInput("y", 
                "Select Y Variable", 
                choices = c("avg_hit_rating", "avg_num_screens", "avg_revenue_inr", "avg_budget_inr"), 
                selected = "avg_hit_rating"),
  
  plotOutput("plot", width="500px", height="400px")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    by_genre %>% 
      ggplot(aes_string(x="genre", y=input$y))+
      geom_col(fill="skyblue", color="black")+
      ggtitle(paste(input$y, "by genre"))+
      theme(plot.title = element_text(size = 16, hjust=0.5))+
      theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9),
            axis.title.x = element_text(size=12),
            axis.title.y = element_text(size=12))
  })
  
}

shinyApp(ui, server)
```

